\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-04-30 ase20/eval.tex
Till Sec. 5.2

\subsection{Check with Tests}\label{sec:check-with-tests}

To evaluate the correctness of \( \tool \), we utilizes Test262, the official
conformance test suite provided by TC39.  First, we complete missing parts of
the AST-IR translator for the latest version of ECMAScript (ES10). Then, we
test the extracted semantics of ES10 with Test262 and reason about the failed
tests.

To focus on core language semantics of JavaScript, we completed only necessary
parts in missing parts of AST-IR translator for ES10.  For abstract algorithms
in ES10, \inred{8,901} steps out of \inred{9,674} steps are automatically
compiled via \textsf{Algorithm Compiler} as described in
Figure~\ref{fig:all-version}(a).  Among the remaining \inred{773} steps, we
manually implemented all missing steps for core language semantics (\inred{221}
steps) and essential parts of built-in libraries (\inred{XX} steps out of
\inred{552}).  Moreover, we manually implemented the \textsf{Global Setting} as
described in Section~\ref{sec:compiler-impl} for the core language features.
However, we do not support several minor language features that do not affect
the overall JavaScript semantics critically: non-strict mode codes, modules,
early errors before actual executions, and inessential built-in objects.

\begin{table*}[t]
  \centering
  \caption{Specification errors in ES10 and in proposals that are ready for
  inclusion in ES11.}
  \label{table:spec-errors}
  \vspace*{-0.5em}
  \small
  \begin{tabular}{|c|c|l|c|c|c|c|c|}
    \hline
    \multicolumn{1}{|c|}{\bf Name} &
    \multicolumn{1}{c|}{\bf Feature} &
    \multicolumn{1}{c|}{\bf Description} &
    \multicolumn{1}{c|}{\bf Known} &
    \multicolumn{1}{c|}{\bf Created} &
    \multicolumn{1}{c|}{\bf Resolved} &
    \multicolumn{1}{c|}{\bf Existed} &
    \multicolumn{1}{c|}{\bf \# Failed Tests} \\\hline

    es10-1 &
    Condition &
    \makecell[l]{Ambiguous production rule for \\
    dangling \code{else} problem in {\it IfStatement}.} &
    \textsf{X} &
    2015-06-01 &
    \inred{2020-00-10} &
    \inred{XXX} days &
    1 \\\hline

    es10-2 &
    Await &
    \makecell[l]{Passing wrong type of \\ the second argument for {\bf PromiseResolve}.} &
    \textsf{O} &
    2019-02-27 &
    2019-04-13 &
    45 days &
    1100 \\\hline

    es10-3 &
    Iteration &
    \makecell[l]{Missing \code{async-iterate} case \\ in the assertion of {\bf
    ForIn/OfHeadEvaluation}.} &
    \textsf{X} &
    2018-02-16 &
    2020-03-25 &
    768 days &
    1120 \\\hline

    es10-4 &
    Completion &
    \makecell[l]{Not handling abrupt completion \\ in {\bf Evaluation} of {\it
    EqualityExpression}.} &
    \textsf{O} &
    2015-06-01 &
    2019-05-02 &
    1065 days &
    1 \\\hline

    es10-5 &
    String &
    \makecell[l]{Wrong use of \code{=} operator in {\bf
    StringGetOwnProperty}.} &
    \textsf{X} &
    2015-06-01 &
    \inred{2019-00-00} &
    \inred{XX} days &
    \inred{XX} \\\hline

    es10-6 &
    Function &
    \makecell[l]{No semantics of {\bf IsFunctionDefinition} \\ for
    \code{function(...)\{...\}}.} &
    \textsf{O} &
    2015-10-30 &
    2020-01-18 &
    1541 days &
    \inred{XX} \\\hline

    es10-7 &
    Function &
    \makecell[l]{No semantics of {\bf ExpectedArgumentCount} \\ for the base case
    of {\it FormalParameters}.} &
    \textsf{O} &
    2016-11-02 &
    2020-02-20 &
    1205 days &
    \inred{XX} \\\hline

    es10-8 &
    Iteration &
    \makecell[l]{Two semantics of {\bf VarScopedDeclarations} \\ for \code{for
    await(var x of e)\{...\}}.} &
    \textsf{O} &
    2018-02-16 &
    2019-10-11 &
    602 days &
    \inred{XX} \\\hline

    bigint-1 &
    Expression &
    \makecell[l]{Using wrong variable \code{oldvalue} instead of \\
    \code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    bigint-2 &
    Number &
    \makecell[l]{Using {\bf ToInt32} instead of {\bf ToUint32} \\ in {\bf
    Number::unsignedRightShift}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    bigint-3 &
    Number &
    \makecell[l]{Using {\bf ToUint32} instead of {\bf ToInt32} \\ in {\bf
    NumberBitwiseOp}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    bigint-4 &
    Number &
    \makecell[l]{Not handling BigInt values in {\bf Number} constructor.} &
    \textsf{O} &
    2019-09-27 &
    2019-11-19 &
    53 days &
    \inred{XX} \\\hline

  \end{tabular}
\end{table*}

To check the correctness of the extracted IR-based semantics using Test262, we
fixed version of Test262 as of February 28, 2019 because the ES10 was branched
out with the branch name es2019 in that day.  While it consists of
\inred{XX,XXX} tests, we filtered out \inred{XX,XXX} tests not applicable for
our formal semantics as shown in Table~\ref{table:test262}.  Test262 targets not
only the current official version of ECMAScript but several extensions and
in-progress features. We excluded \inred{X,XXX} tests for such cases to focus on
the semantics of ES10.  Among \inred{XX,XXX} ES10 tests, we also filtered out
\inred{X,XXX} tests using language features not supported by our semantics.

\begin{table}[t]
  \centering
  \caption{Test results for Test262}
  \label{table:test262}
  \vspace*{-0.5em}
  \small
  \begin{tabular}{lr}\toprule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{All Test262 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Annexes & \inred{XXX}\\\hdashline
    Internationalization & \inred{XXX}\\\hdashline
    In-progress features & \inred{X,XXX}\\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{ES10 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Non-strict modes & \inred{X,XXX}\\\hdashline
    Modules & \inred{X,XXX} \\\hdashline
    Early errors & \inred{X,XXX} \\\hdashline
    Inessential built-in objects & \inred{X,XXX} \\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{Applicable Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Passed tests & \inred{XX,XXX} \\\hdashline
    Failed tests & \inred{XXX} \\\bottomrule
  \end{tabular}
  \vspace*{-2em}
\end{table}

For \inred{XX,XXX} applicable tests, the extracted semantics fails \inred{XXX}
tests.  We reason about the failed tests, and we discover that they failed
because ES10 actually has specification errors. We found eight specification
errors in ES10 (es10-1 to es10-9 in Table~\ref{table:spec-errors}).  Among them,
five errors are already fixed in the draft version of ECMAScript but three errors
are newly revealed.  We reported four new errors, and TC39 confirmed and
merged them into the current draft.

es10-1 is related to the dangling \code{else} problem introduced in ALGOL
60~\cite{dangling-else}. While ES10 describes the rule that the \code{else}
should be associated with the nearest \code{if}, the rule is not explicitly
represented in the ES10 grammar rule.  It causes one test failed and it is a not
yet discovered error.  Thus, we revised the grammar rule to satisfy this rule
and TC39 accepted it in \inred{May X, 2020}.

es10-2 is related to the JavaScript Await.  The {\bf PromiseResolve (C, x)}
algorithm creates a new Promise object using a given constructor {\bf C} and
resolved with a given value {\bf x}.  However, a specification update added
three invocations of {\bf PromiseResolve} with a list of values for {\bf x} in
February 27, 2019.  This error exists in 45 days until April 13, 2019 and causes
\inred{XX} tests failed.

es10-3 arises because of the wrong assertion. ES9 introduced \code{for await}
iteration statements thus it also introduced a new {\it iterationKind} tag,
\code{async-iterate}. However, \code{ForIn/OfHeadEvaluation} algorithm misses
the case of \code{async-iterate} in the assertion. It causes 1120 tests failed
and it is an unknown error. Thus, we revised the assertion to cover the
\code{async-iterate} case and TC39 accepted it in March 25, 2020.

es10-4 and es10-5 arise because contributors misunderstood the meaning of
operations used in abstract algorithms. Especially, es10-5 is related to
\code{=} operator for numbers and not yet discovered. In abstract algorithms,
``{\it x} = {\it y}'' denotes equality expression for double-precision 64-bit
binary format IEEE 754-2008 values thus ``+0 = -0'' is true. However, the {\bf
StringGetOwnProperty} algorithms uses ``{\it index} = -0'' to check whether
index is exactly same with -0 but it is true even if {\it index} is +0.

es10-6, es10-7, and es10-8 are errors that miss or define twice some parts of
cases in semantics. For example, es10-8 is an error that define the {\bf
VarScopedDeclarations} for \code{for await} iterations. If the choice is wrong,
\inred{XX} tests will be failed.

After resolving above specification errors in ES10, we extracted formal
semantics from the revised specification again. The revised formal semantics of
ES10 passed all \inred{XX,XXX} applicable tests in Test262.  \( \tool \) is
helpful to correctly extract semantics from ECMAScript but also to detect bugs
by directly check ECMAScript with Test262. Without our tool, above eight errors
existed in ECMAScript in \inred{XXX} days on average. Besides, three errors
es10-1, es10-3, and es10-5 are even not known before we reported to TC39. We
believe that \( \tool \) bridges gaps between ECMAScript and Test262.

@@

Based on the failed tests, we found three undiscovered specification errors
(bigint-1, bigint-2, and bigint-3), and one reported error (bigint-4) in the
BigInt proposal as described in Table~\ref{table:spec-errors}.  All of
them were confirmed by TC39 and will be fixed in the next release.  The BigInt
proposal added a new type of primitives BigInt and unified the type with
original Number type to Numeric types.  Thus, it not only added new
algorithms for BigInt but also modified all existing algorithms related to
number values.  The error bigint-1 is just a simple misuse of the variable
\code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}. bigint-2 and
bigint-3 break the backward compatibility because of the incorrectly mixed uses
of {\bf ToInt32} and {\bf ToUint32} algorithms in unsigned right shift operators
and bitwise operators, respectively. The last one bigint-4 arises because the
Number constructor does not consider BigInt primitives.  On average, ECMAScript
included these four specification errors in \inred{XXX} days.

>>>>

\subsection{Correctness}\label{sec:check-with-tests}
\begin{table*}[t]
  \centering
  \caption{Specification errors in ES10 and the BigInt proposal ready for
  inclusion in ES11}
  \label{table:spec-errors}
  \vspace*{-.5em}
  \small
  \begin{tabular}{@{}c?c|l|@{}c@{}|c|c|r|@{}r@{}}
\multicolumn{1}{@{}c?}{\bf Name} &
\multicolumn{1}{c}{\bf Feature} &
\multicolumn{1}{c}{\bf Description} &
\multicolumn{1}{@{}c@{}}{\bf Known} &
\multicolumn{1}{c}{\bf Created} &
\multicolumn{1}{c}{\bf Resolved} &
\multicolumn{1}{c}{\bf Existed} &
\multicolumn{1}{@{}c@{}}{\bf \# Fails} \\\toprule
&&&&&&&\\[-1.5em]
    ES10-1 &
    Iteration &
    \makecell[l]{Missing the \code{async-iterate} case in the assertion of\\
 {\bf ForIn/OfHeadEvaluation}} &
    \textsf{X} &
    2018-02-16 &
    2020-03-25 &
    768 days &
    1,120 \\\hline

    ES10-2 &
    Condition &
    \makecell[l]{Ambiguous grammar production for the dangling \code{else} problem in\\
    {\it IfStatement}} &
    \textsf{X} &
    2015-06-01 &
    TBD &
    \multicolumn{1}{c|}{TBD} &
    1 \\\hline

    ES10-3 &
    String &
    \makecell[l]{Wrong use of the \code{=} operator in {\bf
    StringGetOwnProperty}} &
    \textsf{X} &
    2015-06-01 &
    TBD &
    \multicolumn{1}{c|}{TBD} &
    \inred{XX} \\\hline

 ES10-4 &
 Completion &
 Unhandling abrupt completion in {\bf Abstract Equality Comparison} &
 \textsf{X} &
\inred{0000-00-00}&
 2020-04-28 &
 \inred{XX} days &
 \inred{XX}
 \\\hline

    ES10-5 &
    Completion &
    \makecell[l]{Unhandling abrupt completion in {\bf Evaluation} of {\it
    EqualityExpression}} &
    \textsf{O} &
    2015-06-01 &
    2019-05-02 &
    1,065 days &
    1 \\\hline

    ES10-6 &
    Await &
    \makecell[l]{Passing a value of wrong type to the second parameter of\\ {\bf PromiseResolve}} &
    \textsf{O} &
    2019-02-27 &
    2019-04-13 &
    45 days &
    1,100 \\\hline

    ES10-7 &
    Function &
    \makecell[l]{No semantics of {\bf IsFunctionDefinition} for
    \code{function(...)\{...\}}} &
    \textsf{O} &
    2015-10-30 &
    2020-01-18 &
    1,541 days &
    \inred{XX} \\\hline

    ES10-8 &
    Function &
    \makecell[l]{No semantics of {\bf ExpectedArgumentCount} for the base case
    of\\ {\it FormalParameters}} &
    \textsf{O} &
    2016-11-02 &
    2020-02-20 &
    1,205 days &
    \inred{XX} \\\hline

    ES10-9 &
    Iteration &
    \makecell[l]{Two semantics of {\bf VarScopedDeclarations} for\\ \code{for
    await(var x of e)\{...\}}} &
    \textsf{O} &
    2018-02-16 &
    2019-10-11 &
    602 days &
    \inred{XX} \\\hline\hline

    BigInt-1 &
    Expression &
    \makecell[l]{Using wrong variable \code{oldvalue} instead of \\
    \code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-2 &
    Number &
    \makecell[l]{Using {\bf ToInt32} instead of {\bf ToUint32} \\ in {\bf
    Number::unsignedRightShift}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-3 &
    Number &
    \makecell[l]{Using {\bf ToUint32} instead of {\bf ToInt32} in {\bf
    NumberBitwiseOp}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-4 &
    Number &
    \makecell[l]{Not handling BigInt values in {\bf Number} constructor.} &
    \textsf{O} &
    2019-09-27 &
    2019-11-19 &
    53 days &
    \inred{XX}
  \end{tabular}
\end{table*}

\begin{table}[t]
  \centering
  \caption{Test results for Test262}
  \label{table:test262}
  \vspace*{-.5em}
  \small
  \begin{tabular}{lr}\toprule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{All Test262 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Annexes & \inred{XXX}\\\hdashline
    Internationalization & \inred{XXX}\\\hdashline
    In-progress features & \inred{X,XXX}\\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{ES10 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Non-strict mode & \inred{X,XXX}\\\hdashline
    Modules & \inred{X,XXX} \\\hdashline
    Early errors before actual execution & \inred{X,XXX} \\\hdashline
    Inessential built-in objects & \inred{X,XXX} \\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{Applicable Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Passed tests & \inred{XX,XXX} \\\hdashline
    Failed tests & \inred{XXX} \\\bottomrule
  \end{tabular}
  \vspace*{-1em}
\end{table}

To evaluate the correctness of \( \tool \), we checked whether the
extracted semantics from the latest ECMAScript (ES10) passes Test262
as of February 28, 2019, when ES10 was branched out.
To focus on the core language semantics of JavaScript, we completed
only necessary parts missing from the extracted AST-IR translator.
As Figure~\ref{fig:all-version}(a) shows, for the abstract algorithms
in ES10, \inred{8,901} steps out of \inred{9,674} steps are
automatically compiled by \textsf{Algorithm Compiler}.
Among the remaining \inred{773} steps, we manually implemented all
missing steps for the core language semantics (\inred{221} steps) and
the essential parts of the built-in libraries (\inred{XX} steps out of
\inred{552}).  We also manually implemented \textsf{Global Setting}
as described in Section~\ref{sec:compiler-impl} for the core language features.
Note that we do not support minor language features such as the
non-strict mode, modules, early errors before actual execution, and
inessential built-in objects.
Among \inred{XX,XXX} tests in Test262, we filtered out \inred{XX,XXX}
tests as summarized in Table~\ref{table:test262}.  To focus on ES10,
we excluded \inred{X,XXX} tests for annexes, internationalization, and
in-progress features.  We also filted out \inred{X,XXX} tests that use
minor language features.
Finally, for \inred{XX,XXX} applicable tests, the extracted semantics
failed for \inred{XXX} tests.

We investigated the failed tests and found out that they failed due to
specification errors in ES10.  Using the failed tests, we
discovered nine errors: ES10-1 to ES10-9 in
Table~\ref{table:spec-errors}.  Among them, five errors (ES10-5 to
ES10-9) were previously reported and fixed in the current draft of the
next ECMAScript, and the remaining four errors (ES10-1 to ES10-4) were
never reported before.  All four errors were confirmed by TC39, and
will be fixed in the next ECMAScript, ES11.

The specification error ES10-1 is due to a wrong assertion.
While ES9 introduced the \code{for await} iteration statement with a
new {\it iterationKind} tag, \code{async-iterate}, the
\code{ForIn/OfHeadEvaluation} algorithm missed the
\code{async-iterate} case in an assertion, which caused 1,120
tests failed.  We reported the error and proposed a specification fix
to include  the \code{async-iterate} case, and TC39 accepted it on
March 25, 2020.  Because the error was created on February 16, 2018,
it existed for 768 days.

ES10-2 comes from the well-known dangling \code{else} problem
introduced in ALGOL 60~\cite{dangling-else}.  ES10 describes how to
parse it in prose: the \code{else} statement should be associated with
the nearest \code{if} statement.  Because it is written in prose
rather than in the ES10 grammar productions, it caused one test
failed.  We proposed a fix to revise the ambiguous grammar production,
and TC39 confirmed it on April 23, 2020.

ES10-3 is due to a misuse of the \code{=} operator for numbers.
In abstract algorithms, ``{\it x} = {\it y}'' denotes equality testing
for double-precision 64-bit binary format IEEE 754-2008 values;
thus, \mbox{``+0 = -0''} evaluates to true.  However, to check whether
{\it index} is exactly the same with -0,
{\bf StringGetOwnProperty} used \mbox{``{\it index} = -0''},
which is true even when {\it index} is +0.
It caused \inred{XX} tests failed.  We proposed a fix, which is
confirmed on April 22, 2020.

ES10-4 and ES10-5 happened because ES10 did not handle
abrupt completion from function calls.  Our proposed fix to ES10-4 was
accepted on April 28, 2020, and ES10-5 was resolved on May 2, 2019
after existing for 1,065 days.

ES10-6 is due to incorrect uses of an abstract algorithm.
While {\bf PromiseResolve(C, x)} expects a JavaScript object for its
second argument, ES10 passed a list of values rather than an object in
three invocations of {\bf PromiseResolve}.  The wrong invocations were
introduced on February 27, 2019 and caused \inred{XX} tests failed.
They were fixed on April 13, 2019 after existing 45 days.

ES10-7 and ES10-8 happened because ES10 missed semantics for some
cases.  They both existed for more than 1,200 days.

ES10-9 is due to multiple semantics.  Because only one semantics is
intended, \inred{XX} tests fail when a wrong semantics is chosen.

After resolving the nine specification errors in ES10, we extracted a
semantics from the revised specification.  The extracted semantics
from the revised ES10 passed all \inred{XX,XXX} applicable tests in Test262,
which shows that \( \tool \) extracts an IR-based formal semantics
from ECMAScript correctly.  In addition, the evaluation witnesses that
\( \tool \) can detect specification errors effectively.
We could detect not only five previously-known errors
but also four new errors.  We believe that \( \tool \) bridges gaps
between ECMAScript written in a natural language and executable tests
in Test262.

@@

Based on the failed tests, we found three undiscovered specification errors
(BigInt-1, BigInt-2, and BigInt-3), and one reported error (BigInt-4) in the
BigInt proposal as described in Table~\ref{table:spec-errors}.  All of
them were confirmed by TC39 and will be fixed in the next release.  The BigInt
proposal added a new type of primitives BigInt and unified the type with
original Number type to Numeric types.  Thus, it not only added new
algorithms for BigInt but also modified all existing algorithms related to
number values.  The error BigInt-1 is just a simple misuse of the variable
\code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}. BigInt-2 and
BigInt-3 break the backward compatibility because of the incorrectly mixed uses
of {\bf ToInt32} and {\bf ToUint32} algorithms in unsigned right shift operators
and bitwise operators, respectively. The last one BigInt-4 arises because the
Number constructor does not consider BigInt primitives.  On average, ECMAScript
included these four specification errors in \inred{XXX} days.
